%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

Attending university long enough to earn a PhD is an unbelievable privilege, and I only had access to it thanks to an incredible amount of guidance and support.

Most of all, to \DIFaddbegin \DIFadd{my }\DIFaddend mom Calisa Pierce, dad Jim Pierce, \DIFdelbegin \DIFdel{my }\DIFdelend sisters Kelsey and Kim, and \DIFdelbegin \DIFdel{my grandmother Wilma Steele, you've formed all of the best parts of my personality and guided me toward this point, and }\DIFdelend \DIFaddbegin \DIFadd{grandmother Wilma Avis, }\DIFaddend it's impossible to explain how thankful I am for \DIFdelbegin \DIFdel{that. Regarding family I will also always be thankful for }\DIFdelend \DIFaddbegin \DIFadd{your guidance. Thanks also to }\DIFaddend my great aunt Alice who graciously paid for my undergraduate education.

Thanks to my early teachers who put in the extra care to keep me on track, Jonathan Ramey, \DIFaddbegin \DIFadd{Linda Ball, }\DIFaddend Damon Spurlock, Linda Mendez, Linda \DIFdelbegin \DIFdel{Afzhalirhad}\DIFdelend \DIFaddbegin \DIFadd{Afzalirad}\DIFaddend , Phyllis Adkins, Brandon Willard, Marty Ojeda, and Red Hamlin among many others.

Thanks also to my professors at Southern WV Community College who fostered my way of thinking and provided many of the tools I used for the research in this thesis.
At SWVCC, I'd like to thank especially Mindy Saunders, Tex Wood, Anne Klein, Charles Wood, Don Saunders, and George Trimble. \DIFdelbegin \DIFdel{Mindy built my math and physics foundation with her incredibly effective teaching, Anne showed me there's invisible order in everything, and Tex demonstrated that art will always lead science. }\DIFdelend SWVCC was a personal transformation and a great privilege. 

At WVU, I'd like to thank Alan Bristow, Tudor Stanescu, and Leo Golubovic in particular. Alan and Tudor, thank you for providing my first opportunities in research. Leo, you are the key person in my development as a researcher.
\DIFdelbegin \DIFdel{Every time I solve a challenging problem with some Bessel functions or contour integrals, I think ``Dr. Golubovic would like this one!" I'll be successful if I can one day be as influential an educator as you've been.
}\DIFdelend 

To my extended cohort at WVU physics, Payne, Megat, Collins, Scott, Dustin, Rice, Wolfe, Samet, Craig, April, Evan, Larry, Robert, Stephen, Rick, Matt, Caleb, John, and all of you others, thanks for the collaboration and the hilarious times. \DIFdelbegin \DIFdel{I'll never forget Scott's quaternions, Samet's olympiad problems, or Megat's power tower magic (which still blows my mind). I hope the lounge is still happening, that its residents by some miracle still have 24 hour access into the building, that some barefoot Megat-like character is still gaming in the corner, that they have some Andrew Rice imitator for comic relief, and that an esteemed graduate student like Payne still finds time to stoop to the lounge-residents' level. }\DIFdelend May the lounge live on forever!

\DIFdelbegin \DIFdel{Now we come to Marwan Hassan. Some }\DIFdelend \DIFaddbegin \DIFadd{To Marwan Hassan, some }\DIFaddend time ago now I wandered into your office saying I liked rivers and wanted to describe sediment transport with statistical physics. I realize now when I said ``statistical physics" your eyes lit up with visions of Einstein, Nakagawa, Yano, and so many other names I now know well. I was lucky to use the right trigger words.
Marwan, thanks for \DIFdelbegin \DIFdel{taking the risk on me, for }\DIFdelend all of the guidance, for lending me \DIFdelbegin \DIFdel{hundreds }\DIFdelend \DIFaddbegin \DIFadd{many dozens }\DIFaddend of books and papers, for \DIFdelbegin \DIFdel{giving me }\DIFdelend \DIFaddbegin \DIFadd{the }\DIFaddend freedom to explore ideas, and most of all for showing me that it's ok (and implicitly expected) to be human in science. 

My friends in geography have also been great. Thanks to the whole extended research family: Shawn, Matteo, Conor, Nisreen, Yinlue, Alex, Kyle, Leo, Dave, Tobias, Katie, Maria, Elli, Jiamei, Xingyu, Niannian, and Emma, plus the outsider Eaton-ites, Will, Dave, Rose, and Anya. In particular thanks to Shawn for all of the shared speculations\DIFdelbegin \DIFdel{, and Matteo and Conor -- friendships forged through suffering! We'll have to keep in touch and form productive research collaborations once Matteo returns from retirement}\DIFdelend .

Thank you Rui Ferreira and Brett Eaton!
Rui, you welcomed me to Portugal, offered me guidance in the laboratory, and educated me on turbulence and hydraulics. \DIFdelbegin \DIFdel{Your kindness, physics knowledge, and wide-ranging interests beyond physical science are inspiring. Thanks }\DIFdelend \DIFaddbegin \DIFadd{Thanks especially }\DIFaddend for the thoughtful comments \DIFdelbegin \DIFdel{and advice on my PhD. I'm not forgetting what I owe you! I will always be impressed at your concise way of speaking and unbounded kindness}\DIFdelend \DIFaddbegin \DIFadd{on my thesis}\DIFaddend .
Brett, you were the first person to welcome me to UBC, and I'll always appreciate your effort to teach me how to write, interpret flume experiments, and understand simplified \DIFdelbegin \DIFdel{modelling }\DIFdelend \DIFaddbegin \DIFadd{modeling }\DIFaddend in light of real world complexity. 
\DIFdelbegin \DIFdel{Although I'm not sure if I understood your lessons as well as you would have liked, I still have a list of your favorite books on my computer, and they're a first task for as soon as I'm done. Maybe they'll bring me closer to your way of thinking.
}\DIFdelend 

\DIFdelbegin \DIFdel{I'd be amiss not to mention David Furbish. Before }\DIFdelend \DIFaddbegin \DIFadd{To David Furbish, before }\DIFaddend I started my PhD \DIFdelbegin \DIFdel{, }\DIFdelend when I was still \DIFdelbegin \DIFdel{exclusively }\DIFdelend a physics kid \DIFdelbegin \DIFdel{, I attended Furbish's colloquium in Geography at }\DIFdelend \DIFaddbegin \DIFadd{exclusively, I recall attending your colloquium at UBC with }\DIFaddend Marwan's invitation. At the end I asked a vague question about whether Langevin equations might be useful for describing sediment transport, and the answer was ``absolutely yes". Now I have a thesis on it. \DIFdelbegin \DIFdel{David later recruited me into a lovely series of meetings where we discussed gases with adhesion, moon craters, entropy, soft matter, martian hillslopes, the social impacts of disinformation, and a whole range of questions in natural philosophy. David, }\DIFdelend \DIFaddbegin \DIFadd{Thanks for the confirmation! And to the Furbish people -- }\DIFaddend Rachel, Sarah W, Sarah Z, Nakul, Tyler, Erika, and Shawn, thanks sincerely for letting me board the spaceship.

Among UBC Geographers I'd like to thank Nina Hewitt in particular. It was always a pleasure \DIFdelbegin \DIFdel{and sometimes a sigh of relief to TA }\DIFdelend \DIFaddbegin \DIFadd{to work for you on }\DIFaddend your courses, which are so uniquely organized and student-focused. UBC Geography is a much greater department with you in it.

Finally, Mary! My PhD has not been great in terms of work-life balance. Thanks for \DIFdelbegin \DIFdel{sticking it out, }\DIFdelend putting up with my late bedtimes, early \DIFdelbegin \DIFdel{wakeups, the }\DIFdelend \DIFaddbegin \DIFadd{wake-ups, }\DIFaddend rambling science jargon, and \DIFdelbegin \DIFdel{all of the social eventsI skipped}\DIFdelend \DIFaddbegin \DIFadd{skipped social events}\DIFaddend . It's \DIFdelbegin \DIFdel{time! I'm done. I'm ready for an extensive tour of the gap and some beautiful Queensland beaches with your family, featuring SPF110 and an incredibly giant umbrella. I hope to take up still-lives with Marianne's coaching. After that, we're a step closer to settling what colour cat we're gonna get. I'm personally now leaning toward white.
}\DIFdelend \DIFaddbegin \DIFadd{done!
}\DIFaddend 

Completing a PhD looks like a personal achievement, but it's really not.
It's an expression of 30 years of influence and guidance from other people at every stage, my parents, my sisters, my grandmother, and a huge network of teachers and mentors. Thank you all\DIFdelbegin \DIFdel{for every part of it.
}%DIFDELCMD < 

%DIFDELCMD <  %%%
\DIFdelend \DIFaddbegin \DIFadd{!
 }\DIFaddend